---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# roadworksUK

The goal of roadworksUK is to enable you to access, process and visualis data on UK roadworks, particularly Electronic Transfer of Notifications (EToN) records.

## Installation

<!-- You can install the released version of roadworksUK from [CRAN](https://CRAN.R-project.org) with: -->

```{r, echo=FALSE, eval=FALSE}
install.packages("roadworksUK")
```

Install the package hosted on [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("robinlovelace/roadworksUK")
```

Load the package with:

```{r}
library(roadworksUK)
```


## What's in the package?

There are a number of functions for processing EToN records, as shown by:

```{r}
x <- library(help = roadworksUK)
x$info[[2]]
```

The datasets provided by the package include `kent10`, a minimal dataset containing data from Kent, `htdd_ashford`, ~981 records from Ashford, Kent.
The characteristics of these datasets is demonstrated below:

```{r}
data("htdd_ashford")
dim(htdd_ashford) # a larger dataset
htdd_ashford[1:3, 1:5]
```


```{r}
file_path = system.file("extdata", "kent10.csv.gz", package = "roadworksUK")
file.exists(file_path)
htdd_example = rw_import_elgin_htdd(file_path = file_path)
ncol(htdd_example)
nrow(htdd_example) # a small dataset with 10 rows
```

