---
output: github_document
---

[![Build Status](https://travis-ci.org/ITSLeeds/roadworksUK.svg?branch=master)](https://travis-ci.org/ITSLeeds/roadworksUK)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# roadworksUK

The goal of roadworksUK is to enable you to access, process and visualise data on UK roadworks, particularly Electronic Transfer of Notifications (EToN) records.

## Installation

<!-- You can install the released version of roadworksUK from [CRAN](https://CRAN.R-project.org) with: -->

```{r, echo=FALSE, eval=FALSE}
install.packages("roadworksUK")
```

Install the package hosted on [GitHub](https://github.com/) with:

```{r, message=FALSE}
# install.packages("devtools")
devtools::install_github("ITSLeeds/roadworksUK")
```

Load the package with:

```{r}
library(roadworksUK)
```


## What's in the package?

There are a number of functions for processing EToN records, as shown by:

```{r}
x <- library(help = roadworksUK)
x$info[[2]]
```

The datasets provided by the package include `kent10`, a minimal dataset containing data from Kent, `htdd_ashford`, ~981 records from Ashford, Kent.
The characteristics of these datasets is demonstrated below:

```{r}
data("htdd_ashford")
dim(htdd_ashford) # a larger dataset
htdd_ashford[1:3, 1:5]
```


```{r}
file_path = system.file("extdata", "kent10.csv.gz", package = "roadworksUK")
file.exists(file_path)
htdd_example = rw_import_elgin_htdd(file_path = file_path)
ncol(htdd_example)
nrow(htdd_example) # a small dataset with 10 rows
```

## Example: common roadworks in Ashford

This section explores roadworks `htdd_ashford`, a dataset that is provided by the package: it's made available when you load **roadworksUK**.
It relies on the **tidyverse**:

```{r, message=FALSE}
library(tidyverse)
```

A good way to 'get the measure' of potentially large spatio-temporal datasets is to find their size (in MB/GB/TB and number of rows/columns) and their temporal extent (we'll plot its spatial extent soon).
We can find all of these things for the `htdd_ashford` dataset as follows:

```{r}
pryr::object_size(htdd_ashford) # less than 2 MB - good test dataset
nrow(htdd_ashford)
ncol(htdd_ashford)
range(htdd_ashford$e__date_created)
```

A summary table of the contents of this dataset can be generated as follows:

```{r}
htdd_ashford %>% 
  select(id, responsible_org_name, responsible_org_sector, description) %>% 
  slice(1:10) %>% 
  knitr::kable()
```



The following commands can find-out who reports road works in Ashford, using the **dplyr** package (part of the tidyverse):

```{r, message=FALSE}
htdd_ashford %>% 
  group_by(publisher_name) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```

It's mostly Kent County Council. There are a handful of reports by HE in the region also.
Find out who does the work with the following commands (this finds the top 5 organisations, change then `n` parameter to see more organisations):

```{r}
htdd_ashford %>% 
  group_by(responsible_org_name) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  top_n(n = 5, wt = n)
```

Let's take a look at the temporal distribution of roadworks in the example dataset:

```{r plottime1}
plot(htdd_ashford$e__date_created, htdd_ashford$e__duration_days)
```

This distribution is characteristic of roadworks data: it's not usually logged when it begins but after it ends.
A log of actual reporting dates is illustrated in the next plot:

```{r plottime2}
plot(htdd_ashford$e__date_updated, htdd_ashford$e__duration_days)
```

This shows the log comes from a single month (June) in 2018.
We can do more sophisticated plots building on these examples and using packages such as **ggplot2**.
For now, we will move on to plot the spatial extent of the object:

```{r tmapplot2}
library(tmap)
tmap_mode("view")
tm_basemap(server = leaflet::providers$OpenTopoMap) +
  qtm(htdd_ashford$i__location_point)
```



```{r}
library(tidyverse)
htdd_small = htdd_ashford %>% 
  select(description)
```


